Potential competition in alpine bumble bees is mitigated by imperfect trait-matching and subpartitioning within floral morphotypes

### 2021-03-03 code/bias_niche.Rmd
update writing

morph_genus_difference <- niche_overlap_gen %>%
  select(-niche.overlap, -level) %>%
  full_join(niche_overlap_fam) %>%
  select(-niche.overlap, -level) %>%
  full_join(niche_overlap_ktype) %>%
  select(-niche.overlap, -level) %>%
  mutate(tongue.diff = tongue.diff,
         site = factor(site),
         morph.genus.diff = niche.overlap.ktype - niche.overlap.genus)

ggplot(morph_genus_difference, aes(elev.mean, morph.genus.diff)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~tongue.pair)

ggplot(morph_genus_difference, aes(tongue.pair, morph.genus.diff)) +
  geom_boxplot()
```

@@

get_variables(brm_niche10)
conditional_effects(brm_niche10)
mcmc_plot(brm_niche10, pars = c("^r_tongue*"))
mcmc_plot(brm_niche10, pars = c("^r_site*"))
```

@@

## 3.8 Niche overlap ~ elevation
```{r}
# Beta models
gen_gam_niche.Ib <- bam(niche.overlap ~
                            s(sp.pair, bs = "re") +
                            s(elev.mean, by = sp.pair, k = 6, m = 2),
                          method = "fREML",
                          family = betar(eps = 0.001),
                          discrete = TRUE,
                          select = TRUE,
                          data = filter(niche_pool, level == "genus")) %>% getViz()

check.gamViz(gen_gam_niche.Ib)
summary(gen_gam_niche.Ib)
print(plot.gamViz(gen_gam_niche.Ib, allTerms = TRUE), pages = 1)

gen_gam_niche.GI <- bam(niche.overlap ~
                            s(sp.pair, bs = "re") +
                            s(elev.mean, k = 6, m = 2) +
                            s(elev.mean, by = sp.pair, k = 6, m = 1),
                          method = "fREML",
                          family = betar(eps = 0.001),
                          discrete = TRUE,
                          select = TRUE,
                          data = filter(niche_pool, level == "genus")) %>% getViz()

check.gamViz(gen_gam_niche.GI)
summary(gen_gam_niche.GI)
print(plot.gamViz(gen_gam_niche.GI, allTerms = TRUE), pages = 1)

brm_niche01 <- brm(niche.overlap ~ 1 +
                     s(elev.mean, by = sp.pair, k = 10) +
                     (1 | sp.pair),
                   family = "zero_one_inflated_beta",
                   #file = "../output/brm_niche_01",
                   file = "./brm_niche_01",
                   iter = 2000,
                   chains = 4,
                   data = filter(niche_pool, level == "genus"))
```

>>>>

niche_pool_sitemean <- niche_pool %>%
  group_by(site, level, elev.mean) %>%
  summarize(mean.niche.overlap = mean(niche.overlap))

# morph_genus_difference <- niche_overlap_gen %>%
#   select(-niche.overlap, -level) %>%
#   full_join(niche_overlap_fam) %>%
#   select(-niche.overlap, -level) %>%
#   full_join(niche_overlap_ktype) %>%
#   select(-niche.overlap, -level) %>%
#   mutate(tongue.diff = tongue.diff,
#          site = factor(site),
#          morph.genus.diff = niche.overlap.ktype - niche.overlap.genus) 
# 
# ggplot(morph_genus_difference, aes(elev.mean, morph.genus.diff)) +
#   geom_point() +
#   geom_smooth() +
#   facet_wrap(~tongue.pair)
# 
# ggplot(morph_genus_difference, aes(tongue.pair, morph.genus.diff)) +
#   geom_boxplot()
```

@@

get_variables(brm_niche10)
mcmc_plot(brm_niche10, pars = c("^r_tongue*"))
mcmc_plot(brm_niche10, pars = c("^r_site*"))
```

@@

## 3.8 Niche overlap ~ elevation
```{r}
# Pairwise
gen_gam_niche.I <- bam(niche.overlap ~
                            s(sp.pair, bs = "re") +
                            s(site, bs = "re") + # since I have measures repeated over species pairs within each elevation level
                            s(elev.mean, by = sp.pair, k = 8, m = 2),
                          method = "fREML",
                          family = betar(eps = 0.001),
                          discrete = TRUE,
                          select = TRUE,
                          data = filter(niche_pool, level == "genus" &
                                          bb.sp1 %in% c("pasc", "prat", "soro", "telu", "wurf", # dominant species
                                                        "mont", "muci", "pyre", "mend") & # high elevation specialists
                                          bb.sp2 %in% c("pasc", "prat", "soro", "telu", "wurf",
                                                        "mont", "muci", "pyre", "mend"))) %>% getViz()

check.gamViz(gen_gam_niche.I)
summary(gen_gam_niche.I)
plot(sm(gen_gam_niche.I, 9))
plot(sm(gen_gam_niche.I, 12))
plot(sm(gen_gam_niche.I, 14))
plot(sm(gen_gam_niche.I, 15))
plot(sm(gen_gam_niche.I, 16))
plot(sm(gen_gam_niche.I, 32))

# Community mean
ggplot(niche_pool_sitemean, aes(elev.mean, mean.niche.overlap, color = level)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))

mean_niche_gam.I <- gam(mean.niche.overlap ~
                            level + 
                            s(elev.mean, by = level, k = 8, m = 2),
                          method = "REML",
                          family = betar(eps = 0.001),
                          select = TRUE,
                          data = niche_pool_sitemean) %>% getViz()

check.gamViz(mean_niche_gam.I)
summary(mean_niche_gam.I)
print(plot.gamViz(mean_niche_gam.I, allTerms = TRUE), pages = 1)
```

### 2021-03-03 ms_2/ms_2_v2.Rmd
update
update writing

---
title: "Bumble bees exhibit strong but flexible foraging biases through temporal and elevational gradients of floristic turnover"

@@

*Keywords*: optimal foraging, trait matching, pollinator, *Bombus*, alpine

@@

<!-- Open with a discussion of the problem of foraging efficiently in a dynamic foraging environment, the balance of bias and flexibility, etc. -->

Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies (Plowright and Laverty 1984), or indeed by human investigators (Alem et al. 2016; Loukola et al. 2017), renders almost any flower potentially accessible. Even when a gross mismatch between tongue length and corolla depth makes the nectar of a given flower inaccessible by ordinary means, some bumble bee species will resort readily to biting through the base of the flower to “rob” nectar, incidentally making the nectar available to other prospective foragers as well. Flower handling skills are, moreover, socially transmitted, amplifying the plasticity of individual behavior (Leadbeater and Chittka 2008; Loukola et al. 2017). Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the searching/learning/handling costs associated with exploiting it. At an evolutionary level, the innate biases of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards (Raine and Chittka 2007). These innate biases, however, can be modulated or even nullified by patterns of reward learned during foraging (Kugler 1943; Maharaj et al. 2019), which depend not only on intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees (Ishii 2013; Plowright and Laverty 1984; Sowig 1989).

The topic of floral bias in bumble bees has been approached via both laboratory experimentation (e.g. Kugler 1943; Raine and Chittka 2007) and observational field study (e.g. Heinrich 1976; Dave Goulson, Lye, and Darvill 2008). The former approach enables stronger causal inference with respect to the mechanisms of floral choice, but it suffers from all the usual limitations of biological realism associated with laboratory conditions. Observational field studies, however, tend to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance. The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of environmental variation while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species. When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, true bias can be distinguished from the neutral effects of abundance, and variation in bias can be explored along elevational and temporal dimensions of floristic turnover.

In the present study, we focus on the question of bumble bee foraging bias with respect to floral morphology, and the lability of morphological biases across gradients of floristic variation. At the outset, we emphasize the use of the term “bias” in favor of the term “preference,” since the latter is laden with more assumptions and its inference under field conditions is problematic (Williams 2005). By “bias,” we mean simply an empirical disproportionality between the relative abundance of a given floral taxon or morphological class and its proportional visitation. The exact mechanism of this disproportionality --- whether an arbitrary “preference” on the part of the foraging bee, an unmeasured decoupling of relative abundance and rate of reward, an exogenous effect of competition, etc. --- is beyond the scope of our discussion and probably best studied under laboratory conditions where strict factorial control is possible. We focus instead on the empirical phenomenon of bias, its variation, and its morphological correlations.

The meaning of “morphology,” however, is hardly more clear-cut than that of bias. Bumble bees and flowers vary in an array of interrelated morphological parameters, both within and between species, and the selection of relevant and measurable morphological metrics is nontrivial. For bumble bees, morphological variation consists principally (though not exclusively) of variation in tongue length. Tongue length in bumble bees corresponds functionally to corolla depth in flowers, and there is a well-established (though imperfect) association between tongue length and floral choice (e.g. Inouye 1980). Importantly, though, accessing floral reward involves more than just a match between tongue length and corolla depth. Such matching is entirely bypassed by nectar robbers, but even in the case of “legitimate” foraging, access to reward involves learning to manipulate floral structures that, in some floral morphotypes, are challenging irrespective of tongue length (Laverty 1994). For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries (Westerkamp 1997), and the narrowly campanulate flowers of many Ericaceous flora pose a significant barrier to large-bodied visitors like bumble bees (Wallace 1977). Rather than considering various floral morphological traits independently, we rely on the synthesis of Kugler (1970), who developed a typology consisting of ten primary morphotypes and nested subdivision of each. Kugler’s typology was informed by laboratory studies of bumble bee flower choice (Kugler 1943), and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes (Neumayer and Paulus 1999; Schneller et al. 2014).

<!-- Formulate this section more as a list of motivating questions; perhaps end with something that ties back into the more general introduction. -->

How do floral morphotypes compare in terms of total bumble bee visitation?

How does morphological bias vary across bumble bee species and tongue length classes?

Species' morphological biases give rise to patterns of diet (dis)similarity

How are the floral morphotypes distributed through elevation and time in our mountain ecosystem?

For ubiquitously distributed bumble bee species, how does morphological bias vary through elevation and time?

First, we compare the overall rate of bumble bee visitation across floral morphotypes. Then, we parse visitation rates by bumble bee species and explore the functional relationship between floral morphotype and bumble bee tongue length. These biases, together with the background effects of relative abundance, generate emergent patterns of diet (dis)similarity, which we present in the form of ordination analysis. Lastly, we analyze the lability of the morphological biases of individual bumble bee species through elevation and time, and compare these patterns of bias across bumble bee species and within floral morphotypes. The results of these analyses demonstrate both the strength of bumble bee floral bias, arising in part from morphological constraints, and the flexibility of bumble bee floral bias that enables them to forage adaptively in the dynamic floral communities of mountain landscapes.

@@

### Model 5: The variation of morphological bias through elevation and time

To explore the plasticity of floral biases, we first subsetted our data to include only the dominant bumble bee species (B. hortorum, B. pascuorum, B. pratorum, B. soroeensis, B. terrestris/lucorum, and B. wurflenii) and the most visited floral morphotypes (groups 3, 5, 6, and 7). We then constructed a generalized additive model (GAM) in which binary visitation was modeled as a tensor-product smooth of elevation and day-of-year, grouped by floral morphotype. This response was conditioned on linear interaction terms for bumble bee abundance : percent flower cover and percent flower cover : floral morphotype and a random intercept term for bumble bee species : floral morphotype. See Appendix X for the details of model specification and evaluation.

@@

All analyses were conducted in `R` `r citep(citation())`. Bayesian multilevel models were implemented in Stan (Stan Development Team 2021) via the package `brms` `r citep(citation("brms"))`, and visualizations were constructed with the packages `tidybayes` `r citep(citation("tidybayes"))` and `ggdist` `r citep(citation("ggdist"))`. Ordination analysis and permutation testing were performed with the package `vegan` `r citep(citation("vegan"))` and visualized using the packages `ggplot2` `vegan` `r citep(citation("ggplot2"))` and `ggvegan` (Simpson 2019). GAMs were implemented with the package `mgcv` [@Wood_2017] and visualized with the package `mgcViz` [@Fasiolo2018aa]. Data handling and visualization was performed with the `tidyverse` suite of R packages `r citep(citation("gdm"))`. Annotated R code is available in the Supplementary Material.

@@

## Diet similarity

@@

## The lability of bias through elevation and time

While the intraspecific lability of bumble bee foraging biases has been demonstrated under laboratory conditions (Ings, Raine, and Chittka 2009; Maharaj et al. 2019) and in a localized field study (Raine and Chittka 2007), ours is the first study, to our knowledge, to characterize the in situ lability of foraging biases within multiple species through elevational and temporal gradients of floristic turnover. In our analysis, it is evident that morphological biases in bumble bees, insofar as they are captured by the morphotypology used in our study, are not fixed but rather tuned to floristic (and possibly abiotic) context and therefore dynamic through elevation and time. Notably, bumble bee species differed from one another not just in their overall bias toward given morphotypes but also in when and where they were biased toward a given morphotype. This is especially evident in the case of flag flowers, for which positive bias was maximized in mid-season for B. hortorum, B. pascuorum, B. terrestris/lucorum, and B. wurflenii but in early-season for B. pratorum and B. soroeensis. This pattern corresponds to the overall biases of these species toward flag flowers, which were primary resources for B. hortorum, B. pascuorum, B. terrestris/lucorum, and B. wurflenii but only marginal resources for B. pratorum and B. soroeensis.

>>>>

---
title: "Potential competition in alpine bumble bees is mitigated by imperfect trait-matching and subpartitioning within floral morphotypes"

@@

Species coexistence is constrained in part by competition over shared food sources. Bumble bees are classic model organisms for the study of competition and coexistence, and this topic has taken on new urgency as climate-induced range shifts are expected to disrupt competitive relationships between bumble bee species, particularly along the climatic gradients of the mountain slopes that host the world's richest bumble bee communities. We surveyed floral abundance and recorded bumble bee floral visitation patterns along a ~1400 m elevation gradient in German Alps. Using these data, we first ask how floral biases compare across bumble bee species and whether these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype. We then investigate the patterns of diet overlap between bumble bee species that emerge from their respective floral biases, comparing diet overlap at the nested levels of floral morphotype, floral family, and floral genus. Finally, we ask how both community-mean and species-pairwise diet overlap vary with elevation. While tongue length was a strong predictor of bumble bees' floral morphotype biases, species with similar tongue lengths exhibited divergent biases with respect to certain flora morphotypes. This indicates that factors other than tongue length are involved in flower choice, which might explain the coexistence of bumble bee species with similar tongue lengths. Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap. Thus, even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the genus- or species-level subpartitioning of floral resources within morphotypes. This suggests that floral species richness may be as important as overall floral abundance in avoiding the competitive exclusion of bumble bee species. We find that community-mean diet overlap, indicative of the potential for diffuse competition, generally declines with elevation, while the diet overlap for certain species pairs exhibits mid- or high-elevation peaks. The long-term coexistence of mountain bumble bee species may depend on how climate induced range shifts in bumble bees and their floral hosts will alter the these elevational patterns of diet overlap.

*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine

@@

The role of competition in regulating community assembly and species coexistence has long been a central question in ecology. It has taken on new urgency, however, on account of recent climate change, which threatens to disrupt existing patterns of species interactions [@Tylianakis2008-fi], including competition [@Ettinger2017-pk], with consequences for community composition and function. These effects are likely to be most pronounced along the steep climate gradients of mountain slopes, where climate-induced range shifts are expected to introduce novel competitive interactions within trophic levels and cause spatial or phenological mismatches between organisms and their food resources.

Bumble bees (Hymenoptera: *Bombus* spp.) have become classic model organisms for the study of competition and species coexistence due to their broad and overlapping distributions and their gross similarity in morphology, behavior, and dietary niche [@Plowright1984-lc; @Goulson2008-zy]. In this role, however, bumble bees have proven somewhat enigmatic. On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phylogeny (understood to encompass unmeasured trait variation), providing evidence for competition as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk]. Nevertheless, bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy]. Teasing apart the still-obscure mechanisms of bumble bee coexistence and niche partitioning remains a salient problem in basic ecology, but it has also become an urgent question in the context of conservation. Bumble bees are cold-adapted species [@Heinrich1994-cv] associated mainly with mountain habitats, where they serve as the principal pollinators of many alpine plants [@Williams1998-xi; @Goulson2010-lo]. Warming climates have already driven measurable elevational range shifts in bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], and these are predicted to cause declines or even extinction in many bumble bees species [@Soroye2020-zb]. One mechanism by which climate change is expected to affect bumble bees is via the introduction of novel interspecific competition driven by the advance of lowland bumble bee species into the elevational distributions of their highland counterparts. Thus, the speculative question of how competition has acted in the historic assembly of bumble bee communities has become an unfolding process observable over ecological time scales.

All else held equal, species coexistence is favored when intraspecific competition exceeds interspecific competition, and this stabilizing inequality is favored to the extent that species differ in fundamental dietary niche [@Abrams1977-zp; @Chesson2000-aw]. Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs associated with exploiting it. These costs are closely related to the morphological fittedness between bumble bee species and their floral hosts. Bumble bees and flowers vary in an array of interrelated morphological parameters, both within and between species, but for bumble bees, morphological variation consists principally (though not exclusively) of variation in tongue length. Floral handling time depends on the degree of match between bumble bee tongue length and floral corolla depth [@Inouye1980-mv], and there is a well-established (though imperfect) association between tongue length and floral choice. Importantly, though, accessing floral reward involves more than just a match between tongue length and corolla depth, and in some flowers access to floral rewards involves learning to manipulate floral structures that are challenging irrespective of tongue length (Laverty 1994). For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries (Westerkamp 1997), and the narrowly campanulate flowers of many Ericaceous flora pose a significant barrier to large-bodied visitors like bumble bees (Wallace 1977). Thus the dietary niche of a given bumble bee species in a given ecological context is an emergent product of innate floral biases, rewards experienced during foraging, and learned flower handling skills, and all of which contribute to the phenomenon of interspecific niche overlap upon which competition is predicated.

Empirical measurements of niche overlap, however, are inherently equivocal in their interpretation with respect to competition [@Pianka1972-oi; @Pianka1974-qk; @Abrams1980-am]. Low niche overlap between two species, for example, could be interpreted prospectively as an indication that the *potential* for interspecific competition is low, but the same pattern could also be interpreted retrospectively as a mitigating response to high interspecific competition. Which of these interpretations is most plausible hinges on how closely species' observed niches approximate their fundamental niches (within a given resource context). Where observed niches are close to absolute niches, low niche overlap is an indication of low potential for interspecific competition. Any difference between observed and fundamental niches, though, might be attributable to competition-driven niche partitioning, and niche overlap as a reflection of potential competition cannot be distinguished from niche overlap as a response to realized competition. In other words, niche overlap can only be intepreted as an index of *potential* competition if *realized* competition is assumed to be negligible, and this assumption can only be justified on the basis of additional knowledge of the study system under consideration.

Rather than considering various floral morphological traits independently, we rely on the synthesis of Kugler (1970), who developed a typology consisting of ten primary morphotypes and nested subdivision of each. Kugler’s typology was informed by laboratory studies of bumble bee flower choice (Kugler 1943), and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes (Neumayer and Paulus 1999; Schneller et al. 2014).

<!-- 
That one paper [@Zhongming2020-xj] simply equated diet overlap with competition, and I think that is a mistake. But I think we can make a more nuanced case for its use as an index of potential competition. Indeed, Ingolf interpreted diet overlap in exactly this sense in one of his early papers [@Steffan-Dewenter2000-ll]. In fact, there is a slew of papers, from across a wide range of study systems and organisms, that use the term "potential competition" to interpret diet overlap. I think I'm on solid ground.
-->

Importantly, our study system is one of rich floral resources and harsh abiotic conditions, and it is likely that competition is low, or at least transient, and that bumble bees are limited instead by abiotic conditions during the critical period of spring colony founding. Thus, the patterns in our data likely represent floral choice unconstrained by competition and therefore approximate the absolute dietary niches of the species involved (within the context of the studied floral community). On this basis, we argue for the interpretation of niche overlap as an index of *potential* competition rather than a response to *realized* competition, with the caveat this interpretation may misrepresent certain temporal intervals (e.g. early spring colony founding) or elevational bands (e.g. the sparse alpine grasslands above the tree line ecotone, which in our system falls around 1500 m).

We first ask how floral biases compare across bumble bee species whether these patterns can be explained by implicit trait-matching between bumble bee tongue length and floral morphotype. We then investigate the patterns of dietary niche overlap between bumble bee species that emerge from their respective floral biases, comparing niche overlap at the nested levels of floral morphotype, floral family, and floral genus. Finally, we ask how niche overlap varies with elevation. With respect to elevation, we analyze both pairwise niche overlap and community mean niche overlap, since low pairwise niche overlap could conceal a high potential for diffuse competition [@Pianka1974-qk; MacArthur1972-qz]. We discuss the implications of these findings for bumble bee conservation in a changing climate

@@

### Model 5: Niche overlap by level

### Model 6: Genus-level overlap by species-pair

### Model 7: Community-mean genus-level niche overlap by elevation

### Model 8: Pairwise genus-level niche overlap by elevation

@@

All analyses were conducted in `R` `r citep(citation())`. Bayesian multilevel models were implemented in Stan (Stan Development Team 2021) via the package `brms` `r citep(citation("brms"))`, and visualizations were constructed with the packages `tidybayes` `r citep(citation("tidybayes"))` and `ggdist` `r citep(citation("ggdist"))`. Ordination analysis and permutation testing were performed with the package `vegan` `r citep(citation("vegan"))` and visualized using the packages `ggplot2` `vegan` `r citep(citation("ggplot2"))` and `ggvegan` (Simpson 2019). GAMs were implemented with the package `mgcv` [@Wood_2017] and visualized with the package `mgcViz` [@Fasiolo2018aa]. Data handling and visualization was performed with the `tidyverse` suite of R packages `r citep(citation("tidyverse"))`. Annotated R code is available in the Supplementary Material.

@@

## Diet overlap

@@

To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models. A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the structure of regional species pools than in that of local assemblages [@Hanski1982-ie]. In some cases, competition may simply be precluded by a saturating abundance of floral resources [@Williams1989-vr]. Range shifts in bumble bees have already been documented [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], and the introduction or extinction of species from bumble bee communities, with corresponding addition and release of competitive pressure, can alter both the floral visitation patterns of bumble bees and, consequently, the pollination biology of their host plants [@Brosi2013-mu; @Ishii2013-sm].

We see that community-mean diet overlap, indicative of the potential for diffuse competition, generally declines with elevation [@Pianka1974-qk].

"there  is  still  considerable  scope  for  niche  differentiation  among  species  with  tonguesthat  are  near-identical  in  length." @Goulson2008-zy
